The Overlapping Control Volume Finite Element Method for Multiphase Porous Media Flow Modelling \\ Part I: Numerical Formulation

### 2015-01-09 Papers/Prototype_1_Method/article.tex
updates

Finally, the mass matrix ${\mathbf M}$, gradient matrix
${\mathbf C}$ and discretised source term ${\mathbf s}_{u}$ are:
\[\mathbf{M}=\left(
\begin{array}{ccc}
  \bm{M}^{(1,1)} &\bm{0} &\bm{0} \\
  \bm{0}   &\bm{M}^{(2,2)} &\bm{0} \\
  \bm{0} &\bm{0}  &\bm{M}^{(3,3)}   
\end{array}\right),\]
\[
\bm{C}=\left(
\begin{array}{ccc}
  \bm{C}^{(1,1)}&\bm{0}&\bm{0}\\
  \bm{0}&\bm{C}^{(2,2)}&\bm{0}\\
  \bm{0}&\bm{0}&\bm{C}^{(3,3)}\\
\end{array}\right)\quad \text{ and } \;\;
\bm{s}_u=\left(
\begin{array}{ccc}
  \bm{s}^{(1,1)}&\bm{0}&\bm{0}\\
  \bm{0}&\bm{s}^{(2,2)}&\bm{0}\\
  \bm{0}&\bm{0}&\bm{s}^{(3,3)}\\
\end{array}\right),\quad
\]
with blocks defined by:
\begin{eqnarray}
  && \left[{\mathbf M}^{(i,j)}\right]_{p,q} = \int_{\Omega}  Q_{p} \delta_{ij}
{Q}_{q}  dV, \\
  && {\left[\bm{C}^{(i,i)}\right]_{p,q} = \int_{\Omega} Q_{p} \frac{\partial P_{q}}{\partial r_i} dV - \oint_{\Gamma_{E}} \displaystyle\frac{1}{2} Q_{p} n_i^{(E,q)} P_{q} d\Gamma -  \oint_{\Gamma_{\Omega}} \displaystyle Q_{p} n_i P_{q} d\Gamma} \\
  && {\left[\bm{s}_{u}^{(i,i)}\right]_{p} = \int_{\Omega} Q_p} \left[{\mathbf s}_{u}\right]_i dV - \oint_{\Gamma_{\Omega}}\displaystyle Q_{p} n_i p_{bc} d\Gamma.
\end{eqnarray}
Here $\bm{n}=\left(n_{1},n_{2},n_{3}\right)^{T}$ is the outward-pointing normal vector of the domain $\Omega$. $\bm{n}^{(E,q)}$ is the normal to the boundary of element $E$ and is outward-pointing if $P_{q}$ takes support on $E$, and inward-pointing otherwise.

@@

This formulation ensures that for uniform \jrp{This tensor needs a good name} ${\underline{\underline \sigma}}_{\alpha}$ within each element the balance in the \jrp{is this only true in 1D} 1D extended Darcy flow equations is strictly enforced. The saturations are computed in CV space, whereas the permeability tensor $\left(\mathbf{K}\right)$ is assumed piecewise constant in FE space to allow efficient representation of surface parametrised geometries. Thus the tensor %${\underline{\underline \sigma}}_{\alpha}$:
\begin{eqnarray}
  {\underline {\underline \sigma}}_{\alpha} &=& {\underline {\underline \sigma}}_{\alpha} \left[S_{\alpha}\left(\mathbf{r},t\right), \mu_{\alpha}, \mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right), \mathbf{K}\right] \nonumber \\
&=& \frac{S_{\alpha}\mu_{\alpha}}{\mathcal{K}_{{r}_\alpha}}\mathbf{K}^{-1},
\end{eqnarray}
is piecewise constant within the sub-volumes defined through the non-empty intersections between the elements and control volumes (in the pressure discontinuous formulation, these are precisely the control volumes themselves, see Fig.~\ref{fem_cv_represent_a}(c)). The velocity, formed by the action of ${\underline{\underline \sigma}}_{\alpha}^{-1}$ upon the force density $\bm{v}_\alpha$, has an exact polynomial representation on the sub-volumes of one degree lower that of $p$. For example, for a quadratic pressure variation there will be a piecewise linear velocity field, with discontinuities at element and control volume boundaries (Fig.~\ref{fig:darcys_law}). Thus, in 1D there are $2\times 3$ velocity nodal values per element associated with each phase.

>>>>

Finally, the mass matrix ${\mathbf M}$, gradient matrix
${\mathbf C}$ and discretised source term ${\mathbf s}_{u}$ are:
\[\mathbf{M}=\left(
\begin{array}{ccc}
  \bm{M}^{(1,1)} &\bm{0} &\bm{0} \\
  \bm{0}   &\bm{M}^{(2,2)} &\bm{0} \\
  \bm{0} &\bm{0}  &\bm{M}^{(3,3)}   
\end{array}\right),\]
\[
\bm{C}=\left(
\begin{array}{ccc}
  \bm{C}^{(1,1)}&\bm{0}&\bm{0}\\
  \bm{0}&\bm{C}^{(2,2)}&\bm{0}\\
  \bm{0}&\bm{0}&\bm{C}^{(3,3)}\\
\end{array}\right)\quad \text{ and } \;\;
\bm{s}_u=\left(
\begin{array}{ccc}
  \bm{s}^{(1,1)}&\bm{0}&\bm{0}\\
  \bm{0}&\bm{s}^{(2,2)}&\bm{0}\\
  \bm{0}&\bm{0}&\bm{s}^{(3,3)}\\
\end{array}\right),\quad
\]
with blocks defined by:
\begin{eqnarray}
  && \left[{\mathbf M}^{(i,j)}\right]_{p,q} = \int_{\Omega}  Q_{p} \delta_{ij}
{Q}_{q}  dV, \\
  && {\left[\bm{C}^{(i,i)}\right]_{p,q} = \int_{\Omega} Q_{p} \frac{\partial P_{q}}{\partial r_i} dV - \oint_{\Gamma_{E}} \displaystyle\frac{1}{2} Q_{p} n_i^{(E,q)} P_{q} d\Gamma -  \oint_{\Gamma_{\Omega}} \displaystyle Q_{p} n_i P_{q} d\Gamma} \\
  && {\left[\bm{s}_{u}^{(i,i)}\right]_{p} = \int_{\Omega} Q_p} \left[{\mathbf s}_{u}\right]_i dV - \oint_{\Gamma_{\Omega}}\displaystyle Q_{p} n_i p_{bc} d\Gamma.
\end{eqnarray}
Here $\delta_{i,j} = \left[\underline{\underline{\sigma}}_{\alpha}\right]_{i,j}$,  $\bm{n}=\left(n_{1},n_{2},n_{3}\right)^{T}$ is the outward-pointing normal vector of the domain $\Omega$. $\bm{n}^{(E,q)}$ is the normal to the boundary of element $E$ and is outward-pointing if $P_{q}$ takes support on $E$, and inward-pointing otherwise.

@@

%This formulation ensures that for uniform \jrp{This tensor needs a good name -- JG: I can't think on a good name for this -- it's too non-linear to hold any physical meaning. Maybe solid-fluid homogeneisation factor/ parameter/ coeff} ${\underline{\underline \sigma}}_{\alpha}$ within each element the balance in the \jrp{is this only true in 1D} 1D extended Darcy flow equations is strictly enforced. 
The saturations are computed in CV space, whereas the permeability tensor $\left(\mathbf{K}\right)$ is assumed piecewise constant in FE space to allow efficient representation of surface parametrised geometries. Thus the viscous-friction damping tensor \jrp{This tensor needs a good name JG: I can't think on a good name for this -- it's too non-linear to hold any physical meaning. Maybe solid-fluid homogeneisation factor/ parameter/ coeff or viscous-friction damping ...}, %${\underline{\underline \sigma}}_{\alpha}$:
\begin{eqnarray}
  {\underline {\underline \sigma}}_{\alpha} &=& {\underline {\underline \sigma}}_{\alpha} \left[S_{\alpha}\left(\mathbf{r},t\right), \mu_{\alpha}, \mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right), \mathbf{K}\right] \nonumber \\
&=& \frac{S_{\alpha}\mu_{\alpha}}{\mathcal{K}_{{r}_\alpha}}\mathbf{K}^{-1},
\end{eqnarray}
is piecewise constant within the sub-volumes defined through the non-empty intersections between the elements and control volumes; in the pressure discontinuous formulation, these are precisely the control volumes themselves, Fig.~\ref{fem_cv_represent_a}(c). The velocity, formed by the action of ${\underline{\underline \sigma}}_{\alpha}^{-1}$ upon the force density $\bm{v}_\alpha$, has an exact polynomial representation on the sub-volumes of one degree lower that of $p$. For example, for a quadratic pressure variation there will be a piecewise linear velocity field, with discontinuities at element and control volume boundaries (Fig.~\ref{fig:darcys_law}). Thus, in 1D there are $2\times 3$ velocity nodal values per element associated with each phase.
